Compositional equivalences based on Open pNets

### 2020-01-30 2016-Weak-FH-Bisim/WeakBisim.tex
papier au format lmcs
adapt new sty, et remove \TODO pages 12-13

\documentclass[preprint]{elsarticle}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{prop}{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{example}{Example}
%\newtheorem{problem}[theorem]{Problem}
\newenvironment{proof}{\paragraph{Proof}}{}

\title{Compositional equivalences based on Open pNets\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{Ludovic Henrio  \ \ \  Eric Madelaine \ \ \ Rab\'ea}
%\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
%	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
%}
\date{}                                           % Activate to display a given date or no date

                          % Activate to display a given date or no date

@@

@@

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2008,AmeurBoulifa2017,HKM-FASE16}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example \TODO{TODO}
\end{itemize}

@@

\LUDO{I added this paragraph}
Because of the semantic interpretation of open automata, the set of open transition of an open automaton is infinite (for example because every free variable can be renamed). However an open automaton is characterized by a  subset of these open transition which is sufficient to generate, by substitution the other ones. In the following, we will abusively write that we define an ``open automaton'' when we provide only the set of open transitions that is sufficient to generate a proper open automaton by saturating each open transition by all possible substitutions.

@@

\TODO{J'ai un probleme ici, comme on a dit que les AO etaient satures par substitution, ils sont par nature infinis... Donc le theoreme, tel qu'il est exprime, est vrai, mais inapplicable. La solution est dans le papier de PEPM, mais complique a expliquer...}
\TODO{pour faire simple, on vire le theoreme, et on met une simple note disant qu'on a une definition dans PEPM d'une description structurelle finie, et un algo avec terminaison ?}
\TODO{[LUDO->ERIC] D'accord: tu ecris comme tu connais le mieux le papier PEPM?}
If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation, provided the logic of the predicates is decidable (see \cite{henrio:Forte2016}).

@@

This section defines the semantics of an open pNet as a translation into an open automaton. 
In this translation, the states of the open automata are obtained from the states of the pLTSs at the leaves of the composition. The predicates on the transitions are obtained both from the predicates on the pLTSs transitions and from the synchronisation vectors involved in the transition.
\TODO{Meme remarque ici, on ne genere qu'une representation finie des OAs}

>>>>

\documentclass{lmcs}

@@

\newtheorem{theorem}{Theorem}
%\newtheorem{prop}{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{example}{Example}
%\newtheorem{problem}[theorem]{Problem}
%\newenvironment{proof}{\paragraph{Proof}}{}

@@

\title{Compositional equivalences based on Open pNets
%\thanks{This work was partially funded by the Associated Team FM4CPS
%  between INRIA and ECNU, Shanghai}
}
\author{Ludovic Henrio}

\author{Eric Madelaine}
\address{INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France}
%\date{}                                           % Activate to display a given date or no date
\author{Rab\'ea Ameur-Boulifa}
                          % Activate to display a given date or no date

@@

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2008,AmeurBoulifa2017,henrio:hal-01303557}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example \TODO{TODO}
\end{itemize}

@@

%\LUDO{I added this paragraph}
Because of the semantic interpretation of open automata, the set of open transition of an open automaton is infinite (for example because every free variable can be renamed). However an open automaton is characterized by a  subset of these open transition which is sufficient to generate, by substitution the other ones. In the following, we will abusively write that we define an ``open automaton'' when we provide only the set of open transitions that is sufficient to generate a proper open automaton by saturating each open transition by all possible substitutions.

@@

%\TODO{J'ai un probleme ici, comme on a dit que les AO etaient satures par substitution, ils sont par nature infinis... Donc le theoreme, tel qu'il est exprime, est vrai, mais inapplicable. La solution est dans le papier de PEPM, mais complique a expliquer...}
%\TODO{pour faire simple, on vire le theoreme, et on met une simple note disant qu'on a une definition dans PEPM d'une description structurelle finie, et un algo avec terminaison ?}
%\TODO{[LUDO->ERIC] D'accord: tu ecris comme tu connais le mieux le papier PEPM?}

\paragraph{Note on finite versus infinite open automata:} as
mentionned in page \pageref{def:open-automaton}, we have defined two flavors of
automata. More precisely, in \cite{hou:hal-02406098}, we formely define
\emph{semantic open automata} (infinite as in Definition \ref{def:open-automaton}),
and \emph{structural open automata} (finite) that can be generated as
the semantics of pNets, and used in the implementation. Then we define
an alternative version of our bisimulation, called
Structural-FH-Bisimulation, based on structural open automata, and
prove that the \emph{semantic} and \emph{structural} FH-Bisimulations coincide.
In the sequel, all mentions of finite automata, and algorithms for
bisimulations, implicitely refer to their \emph{structural} versions.

If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation, provided the logic of the predicates is
decidable (see \cite{henrio:Forte2016}). Formaly:

@@

This section defines the semantics of an open pNet as a translation into an open automaton. 
In this translation, the states of the open automata are obtained from
the states of the pLTSs at the leaves of the composition. The
predicates on the transitions are obtained both from the predicates on
the pLTSs transitions and from the synchronisation vectors involved in
the transition.
